### 2016-11-15 response.tex
Before the corrections by Simon

% Created 2016-11-15 Tue 14:14
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{listings}
\date{}
\title{Response to the Referee\\\medskip
\large Article DX11551}
\hypersetup{
 pdfauthor={Oscar Castillo-Felisola},
 pdftitle={Response to the Referee},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 24.4.1 (Org mode 9.0)}, 
 pdflang={English}}
\begin{document}

\maketitle
Dear Referee,

First of all we want to thank your review, constructive comments and
suggestions.

Below we present a list of changes to the manuscript that will clarify
your concerns about our work.

\begin{itemize}
\item Differences four-dimensional approach vs higher dimensional: We have
added extra paragraphs to our discussion section pointing to the
following differences
\begin{itemize}
\item Gravitational models including higher order curvature/torsion are in
general not minimal, whereas our model is (in the sense that
torsion enters only through (non-Riemannian) curvature, and only an
extra coupling constant is introduced in our analysis).
\item The initial model assures that the field equation are first order.
\item The different behavior of the cosmological model are driven by the
relation between the coupling constant solely.
\item All the cosmological scenarios presented are free of \emph{finite time}
future singularities.
\item Some scenarios present dynamical compactification.
\item The scale factor does not depend on the \(k\) parameter which module
the metric structure.
\item The equations of our model is possesses only \emph{CP} even terms.
\end{itemize}

\item We thank the comment about the generalized theories with no ghosts
and second order field equations. On that respect, we have added a
comment in the manuscript, with its respective reference.

\item We have clarified that some of the wormhole solutions, mentioned in
page 2, violate no energy conditions.

\item Your question about future singularities was addressed in the
previous point regarding the difference between four- and five-
dimensional models.
\end{itemize}

\begin{center}
Best regards,

Simon del Pino

On behalf of the collaborators
\end{center}
\end{document}